### 2025-05-17 draft.tex
feat: Add validation of cp_measure vs cprof

% Created 2025-05-16 Fri 08:56
% Intended LaTeX compiler: pdflatex
\documentclass{article}

@@

\begin{figure}[htbp]
\centering
\includesvg[width=.9\linewidth]{./figs/astro3d}
\caption{\label{fig:org7fe862f}{[}Placeholder] Comparison of CellProfiler and cp\_measure.}
\end{figure}
\subsection{Results and examples of usage}
\label{sec:org8ef524e}
\subsubsection{Classification of astrocytes and their distinctive features}
\label{sec:org3d97e6f}

@@

\begin{figure}[htbp]
\centering
\includesvg[width=.9\linewidth]{./figs/shap}
\caption{\label{fig:org4620d4d}Shapley values of most important features to classify the day in which an image was taken (a multi-class classification task). The test data accuracy is shown in bold. Our results showcased the axis length of the cell to be a major indicator of phenotypic effect, implying that cells became more elongated on their minor axis.}
\end{figure}
\subsubsection{Applicability on spatial transcriptomics}
\label{sec:org1d0d8b8}
A key advantage of providing these CellProfiler-like measurements as a standalone Python package is their ease of integration into diverse analytical workflows, which otherwise would require substantial adaptation to the standard CellProfiler environment. The recent proliferation of black-box foundation models trained solely on morphological data highlights morphology as a highly informative and predictive modality. However, the feature vectors produced by these models are typically not interpretable, preventing direct biological assessment. In contrast, classical morphological measurements yield explicit, interpretable readouts—for instance, the co-localization of fluorescent markers—facilitating clear biological interpretations.

>>>>

% Created 2025-05-17 Sat 12:54
% Intended LaTeX compiler: pdflatex
\documentclass{article}

@@

\begin{figure}[htbp]
\centering
\includesvg[width=.9\linewidth]{./figs/jump_r2_examples}
\caption{\label{fig:org7fe862f}{[}Placeholder] Comparison of CellProfiler and cp\_measure.}
\end{figure}
\subsection{Results and examples of usage}
\label{sec:org8ef524e}
\subsubsection{Classification of astrocytes and their distinctive features}
\label{sec:org3d97e6f}

@@

\begin{figure}[htbp]
\centering
\includesvg[width=.9\linewidth]{./figs/example_shap}
\caption{\label{fig:org4620d4d}Shapley values of most important features to classify the day in which an image was taken (a multi-class classification task). The test data accuracy is shown in bold. Our results showcased the axis length of the cell to be a major indicator of phenotypic effect, implying that cells became more elongated on their minor axis.}
\end{figure}
\subsubsection{Applicability on spatial transcriptomics}
\label{sec:org1d0d8b8}
A key advantage of providing these CellProfiler measurements as a standalone Python package is their ease of integration into diverse analytical workflows, which otherwise would require substantial adaptation to the standard CellProfiler environment. The recent proliferation of black-box foundation models trained solely on morphological data highlights morphology as a highly informative and predictive modality. However, the feature vectors produced by these models are typically not interpretable, preventing direct biological assessment. In contrast, classical morphological measurements yield explicit, interpretable readouts—for instance, the co-localization of fluorescent markers—facilitating clear biological interpretations.
